Assigning Scouts to Optimal Patrols

### 2016-05-25 draft1.tex
changes to paper

\begin{abstract}
Scoutmaster Gene Bruce of the Boy Scouts of America for Troop 407 in Kent, Washington wants to find an arrangement of 13 new scouts into two patrols. The cohesiveness and overall quality experience of a patrol is affected by adverse relationships and existing friendships. Thus, by using anonymous data from scouts and parents, a brute force strategy was employed to find the optimal arrangement of scouts that avoids severe conflict and maximizes existing preferences. We found an optimal arrangement of the 13 boy scouts that satisfies the desired conditions.
\end{abstract}
\maketitle
\section*{Problem Description}
Every year the Scoutmaster must assign new scouts to groups called patrols. In these patrols, scouts would plan and perform different scouting activities together in which they can form close friends that grow together. However, in recent years, the number of scouts have been increasing at a rate such that it is no longer logistically viable to place every new scout in a single patrol; multiple patrols are needed to organize activities better and to form proper bonds and community between the new recruits. Furthermore, the new scouts are in the fifth or sixth grade who have not fully developed their interpersonal skills. Many can be temperamental and awkward when dealing with adversity, and even some devastating conflicts occur where not only the two fighting scouts create animosity between each other, the rest of the troop alienates the destructive pair. If enough of these adverse interactions occur within a troop, many scouts would decide to quit and lose out in a great experience and community in the Boy Scouts.

The data the Scoutmaster collects about adverse pairings and preferences are either privately given by the scouts themselves. Possibly destructive interactions are given by the parents. In the past, the Scoutmaster determined the perfect arrangement of patrols without too much time and effort due to the smaller number of new recruits. However, newer scouts are joining yearly to the point that solving this grouping problem cannot be done by hand in a reasonable time. Therefore, we can help our community partner find a better means to solve his troop organization problem. Our goal for this project is to assign boy scouts in different patrols of appropriate size (of 6-8 scouts) such that we maximize the retention rate of the scouts in the program by avoiding severe conflicts while maximizing positive relationships between scouts.

We aim to answer the direct problem our community partner provided of arranging 13 scouts into two troops that avoids severe conflicts and promotes the positive relationships between the scouts. Furthermore, we want to determine if we can build a model to predict these patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years. Additionally, this model can possibly be applied to other situations in which arranging a number of people into groups where interpersonal dynamics is necessary and important such as project teams in a company and forming new military troops.

\section*{Data}
The 13 different recruited scouts are Brandon, Cameron, Christian, Colby, Daniel, Darwin, Evan, Jake, Jordan, Nathan, Patrick, Timmy and Tommy. After the Scoutmaster collected the likes and dislikes of each scout, he presented the data as follows: 
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.60]{graph}
\end{figure}

The solid arrow represents a preferred companion. For example, there is a solid arrow from Tommy pointing to Colby indicating that Tommy likes to be with Colby. A dashed arrow represents a dislike. For example, there is a dashed arrow from Tommy to Jordan showing that Tommy does not want to be with Jordan. Furthermore, the Scoutmaster provided us a separate email, and indicated that Colby and Jordan cannot be in the same troop. The highlighted arrows in the figure above does not mean anything new between the relationships of the boy scouts.

The Stirling number of the second kind provides the total possible ways to arrange any number of distinguishable objects to indistinguishable groups as long as any troop size is permitted and no possible empty troops. For the data in this project, the Stirling number of the second kind for 13 scouts into 2 troops is 4095. This number provides an upper bound on the possible arrangements for the specific problem tackled in this project.

@@

\begin{table}[h]
	\caption{}\label{eqtable}
	\renewcommand\arraystretch{1.5}
	\noindent\[
	\begin{array}{|c|c|}
	\hline
	\textbf{Patrol 1}&\textbf{Patrol 2}\\
	\hline
	\text{Brandon, Cameron, Colby,}&\text{Christian, Daniel, Jake,}\\
	\text{Darwin, Evan, Tommy}&\text{Jordan, Nathan, Patrick, Timmy}\\
	\hline
	\end{array}
	\]
\end{table}

@@

\section*{Appendix B}
The solution for finding the Stirling number of the second kinds for $n = 13$ scouts to be arranged in $k = 2$ troops:
\begin{gather*}
	S(n,k) = \frac{1}{k!} \sum_{i=0}^{k}(-1)^k {2 \choose i} (k - i)^{n} \\
	S(13,2) = \frac{1}{2!} \sum_{i=0}^{2}(-1)^2 {2 \choose i} (2 - i)^{13} \\
	S(13, 2) = 4095
\end{gather*}

@@

\section*{Mathematical Model}
This problem visually lends itself well to the notion of a mathematical graph. Since there is an inherent sense of relationships between individuals (how much a scout likes another scout), it is appropriate to consider the scouts as vertices in a graph with directed edges representing how much the “head” scout likes the “tail” scout. Our goal, using this language, is to find a partition of the vertices in the graph that satisfies some sort of objective function given size constraints of the partitions. In our problem at hand, we consider a complete graph with 13 vertices and edge weights being either 0, 1, or -1, corresponding to indifference, approval, or disapproval respectively of the “head” scout with respect to the “tail” scout. The Scoutmaster specifically wants two groups out of these 13, and with the size constraints for a patrol considered, our goal becomes finding 6 vertices to be in one partition and the other 7 being in another. The total number of partitions becomes 13 choose 6, which is equivalent to 1,716. This is certainly within the scope of considering total enumeration of the partitions as a reasonable solution to the problem, therefore we will proceed accordingly and implement the calculation using Python and its associated libraries to represent the data in a format that is convenient to work with.

@@

\section*{Conclusions}
We initially struggled deciding on an appropriate objective function to partition the scouts to different groups. Follow up interviews with Scoutmaster Gene Bruce allowed us to decide on this function. We relied on his expertise and intuition rather than trying to blindly decide on which types of relationships should be fostered and avoided to determine our objective function. Thus, we learned necessary communication skills and consideration of our community partner's needs.

With a feasibility check of how many solutions possible once the mathematical model was decided using graphs, a brute force strategy was the most reliable method to arrive at a solution. Thus we were able to develop a program that can partition scouts using this strategy. We learned to utilize tools we have at hand. Further consideration of other methods later on to reduce the computational time of our algorithm.

\section*{Acknowledgments}
We would like to thank Scoutmaster Gene Bruce for providing us the necessary data and being extremely patient and understanding when discussing the background and problem. We would also like to thank Professor Sara Billey for the mentorship and guidance she provided throughout this project, and our TA Austin Tran for his comments and suggestions.

>>>>

@@

\begin{table}[h]
	\caption{}\label{eqtable}
	\renewcommand\arraystretch{1.5}
	\noindent\[
	\begin{array}{|c|c|}
	\hline
	\textbf{Patrol 1}&\textbf{Patrol 2}\\
	\hline
	\text{Brandon, Cameron, Colby,}&\text{Christian, Daniel, Jake,}\\
	\text{Darwin, Evan, Tommy}&\text{Jordan, Nathan, Patrick, Timmy}\\
	\hline
	\end{array}
	\]
\end{table}
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.60]{graph}
\end{figure}

@@

@@

\section*{Mathematical Model}
The problem of placing individuals into groups visually lends itself well to the notion of a mathematical graph. In mathematical vocabulary, a graph consists of a group of objects called vertices that get connected with an "edge" if they share some sort of relationship. The graph we will use to represent the data of our problem can be visualized with the image offered on page (5?), where vertices are represented by the circled names and edges are represented by lines between the circles. Since there is an inherent sense of relationships between individuals (how much a scout likes another scout), it is appropriate to consider the scouts as vertices in a graph with directed edges representing how much the “head” scout likes the “tail” scout. Our goal, using this language, is to find a partition of the vertices in the graph that satisfies some sort of objective function given size constraints of the partitions. In our problem at hand, we consider a complete graph with 13 vertices and edge weights being either 0, 1, or -1, corresponding to indifference, approval, or disapproval respectively of the “head” scout with respect to the “tail” scout. The Scoutmaster specifically wants two groups out of these 13, and with the size constraints for a patrol considered, our goal becomes finding 6 vertices to be in one partition and the other 7 being in another. The total number of partitions becomes 13 choose 6, which is equivalent to 1,716. This is certainly within the scope of considering total enumeration of the partitions as a reasonable solution to the problem, therefore we will proceed accordingly and implement the calculation using Python and its associated libraries to represent the data in a format that is convenient to work with.

@@
